Multiphysics Core Dynamics Simulation Using the Improved Quasi-Static Method

### 2017-02-16 doc/M&C_2017/FullPaper/MC2017_FullPaper.tex
Adding more results

%%%% Abstract
\begin{document}
\vspace*{-42pt}
\begin{strip}
\centering{\parbox{153mm}{{\bf Abstract} \itshape - 
The purpose of this paper is to present a new technique in dealing with multi-physics feedback in transient reactor simulations, specifically utilizing the improved quasi-static method (IQS).  the technique involves the addition of an intermediate time scale to evaluate temperature and PRKE parameters.  The theory is that temperature and the parameters are more quickly varying in time than the shape, thus further reducing the number of multi-group diffusion evaluations and execution time for a given accuracy. The technique was implemented into the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) and tested with the LRA benchmark and the Transient-15 TREAT problem.  The results proved that the addition of this time scale was effective in reducing error and execution time for these problems. 
}\par}
\vspace*{14pt}
\end{strip}

@@

\begin{table*}[htb]
\begin{center}
%\resizebox{\linewidth}{!}{
\begin{tabular}{|l|ccccccc|}
\hline
 & No.	& Max &	Time at Max &	Max Average &	\% Increase &	Max Power & Linear	\\
Method & of Steps &	Power (W) &	Power (s) &	Temperature (K) & Runtime$^*$ & Error & Iterations \\
\hline
Baseline 			& 3000 	& 3.5039e+08 & 1.901 & 371 &	---		&	---		 & ---		\\
Brute Force 		& 300 	& 3.5011e+08 & 1.90  & 371 &	---		&	7.875e-4 & 41020	\\
IQS 				& 300	& 3.5036e+08 & 1.90  & 371 & -11.9\%	&	8.385e-5 & 23949	\\
IQS (5 updates) 	& 300 	& 3.5040e+08 & 1.90  & 371 &  49.7\%	&	3.687e-5 & 24035	\\
IQS P-C 			& 66 	& 4.3400e+08 & 1.90  & 374 &  20.3\%	&	0.238	 & 12318	\\
IQS P-C (5 updates) & 66 	& 3.5370e+08 & 1.90  & 372 &  81.9\%	&	0.00935	 & 15598	\\
\hline
\end{tabular}
%}
\end{center}
\vspace{-3mm}
$^*$ difference in runtime from brute force 
\caption{Transient-15 Error and Runtime Results}
\label{tab:tran15}
\end{table*}

The results from \tbl{tab:tran15} show similar performance of IQS with the temperature updates as the LRA.  IQS with 1 temperature update shows a performance that reduces the error to approximately a tenth of the brute force error, and reduces the execution time by about 12\%.  This shows that IQS was able to resolve the nonlinearity between flux and temperature with significantly less diffusion evaluations.  Having IQS with 5 updates significantly increased the execution time for the same time step, but the error was reduced.  Comparing this error to a similar brute force error at a smaller time step could show that the runtime was reduced.  However, IQS P-C did not perform nearly as well and proved almost detrimental to the problem.  With only one update, IQS P-C achieved a less accurate solution and increased the execution time.  Only with 5 updates was the error reduced from the brute force error, but the increase in run-time was significant. The transient profile of the varibles' dynamical time scales is shown in \fig{fig:tran15tc}.  This plot exhibits a similar response to that of the LRA. The response of temperature shows that the updates are a computational frugal treatment of the feedback and adaptation of the number of updates is vital for optimization. \\

@@

Two examples were chosen to test the implementation of the temperature feedback treatment: the LRA benchmark and the Transient-15 TREAT problem.  Both problems involve adiabatic heat up of the core, the LRA has a constant specific heat, while Transient-15's is dependent on temperature.  The LRA showed clear results concluding a marginal improvement in performance for IQS and a impressive performance for IQS P-C.  The results proved that the updates improved the performance for both IQS and IQS P-C, but optimization of the number of updates is important, possibly requiring adaptation.  The implementation to the Transient-15 problem gave more ambiguous results.  IQS proved to significantly benefit from the updates, increasing accuracy and decreasing execution time.  IQS P-C was able to reduce the error with the updates, but the execution time was significantly greater.

In conclusion, the multi-physics treatment with IQS presented in this paper is worth further investigation, optimization, and implementation into other transient reactor problems with other feedback mechanisms.

>>>>

%%%% Abstract
\begin{document}
\vspace*{-42pt}
\begin{strip}
\centering{\parbox{153mm}{{\bf Abstract} \itshape - 
The purpose of this paper is to present a new technique in dealing with multi-physics feedback in transient reactor simulations, specifically utilizing the improved quasi-static method (IQS).  the technique involves the addition of an intermediate time scale to evaluate temperature and PRKE parameters.  The theory is that temperature and the parameters are more quickly varying in time than the shape, thus further reducing the number of multi-group diffusion evaluations and execution time for a given accuracy. The technique was implemented into the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) and tested with the LRA benchmark and the Transient-15 TREAT problem.  The results proved that the addition of this time scale was effective in reducing execution time for these problems while maintaining accuracy. 
}\par}
\vspace*{14pt}
\end{strip}

@@

\begin{table*}[htb]
\begin{center}
%\resizebox{\linewidth}{!}{
\begin{tabular}{|l|ccccccc|}
\hline
 & No.	& Max &	Time at Max &	Max Average &	\% Increase &	Max Power & Linear	\\
Method & of Steps &	Power (W) &	Power (s) &	Temperature (K) & Runtime$^*$ & Error & Iterations \\
\hline
Baseline 			& 3000 	& 3.5039e+08 & 1.901 & 371 &	---		&	---		 & ---		\\
Brute Force 		& 300 	& 3.5011e+08 & 1.90  & 371 &	---		&	7.875e-4 & 41020	\\
IQS 				& 300	& 3.5036e+08 & 1.90  & 371 & -11.9\%	&	8.385e-5 & 23949	\\
IQS (5 updates) 	& 300 	& 3.5040e+08 & 1.90  & 371 &  49.7\%	&	3.687e-5 & 24035	\\
IQS P-C 			& 300 	& 3.5065e+08 & 1.90  & 371 &  -2.1\%	&	7.527e-4 & 39020	\\
IQS P-C (5 updates) & 300 	& 3.5043e+08 & 1.90  & 371 &  26.5\%	&	1.227e-4 & 37866	\\
\hline
\end{tabular}
%}
\end{center}
\vspace{-3mm}
$^*$ difference in runtime from brute force 
\caption{Transient-15 Error and Runtime Results}
\label{tab:tran15}
\end{table*}

The results from \tbl{tab:tran15} show similar performance of IQS with the temperature updates as the LRA.  IQS with 1 temperature update shows a performance that reduces the error to approximately a tenth of the brute force error, and reduces the execution time by about 12\%.  This shows that IQS was able to resolve the nonlinearity between flux and temperature with significantly less diffusion evaluations.  Having IQS with 5 updates significantly increased the execution time for the same time step, but the error was reduced.  Comparing this error to a similar brute force error at a smaller time step could show that the runtime was reduced. IQS P-C performed not nearly as well as it did with the LRA benchmark, but still proved to be effective.  Having 5 updates for IQS P-C increased the runtime marginally, but decreased the error significantly.  The transient profile of the varibles' dynamical time scales is shown in \fig{fig:tran15tc}.  This plot exhibits a similar response to that of the LRA. The response of temperature shows that the updates are a computational frugal treatment of the feedback and adaptation of the number of updates is vital for optimization. \\

@@

Two examples were chosen to test the implementation of the temperature feedback treatment: the LRA benchmark and the Transient-15 TREAT problem.  Both problems involve adiabatic heat up of the core, the LRA has a constant specific heat, while Transient-15's is dependent on temperature.  The LRA showed clear results concluding a marginal improvement in performance for IQS and a impressive performance for IQS P-C.  The results proved that the updates improved the performance for both IQS and IQS P-C, but optimization of the number of updates is important, possibly requiring adaptation.  The implementation to the Transient-15 problem gave similar results.  IQS proved to marginally benefit from the updates, decreasing error but increasing execution time.  IQS P-C was benefited more by the updates in this example. The updates were able to reduce the error with a marginal increase in execution time.

In conclusion, the multi-physics treatment with IQS presented in this paper is worth further investigation, optimization, and implementation into other transient reactor problems with more feedback mechanisms.
